\startsection{Matrices}

Formatting: a $n \times m$ matrix is a matrix with $n$ rows and $m$ columns.

Terminology:\\
Square matrix: a $n \times n$ matrix

A matrix \textit{mod} m is just a matrix in which every number has been \textit{mod}ded by m. When doing this with matrix multiplication, add \inline{sum \%= m;} after line \#6.

The identity matrix is a $n \times n$ matrix where the elements on the main diagonal (top-left to bottom-right) are ones and the rest are zeroes. It is similar to the number 1 in numerical mathematics. 

\subsection*{Addition and subtraction}

Add/subtract each number to its pair in the second matrix. (Note: the matrices must be the same size.)

\begin{equation*}
    \begin{bmatrix}
        2 & 6 & 12 \\
        7 & -4 & 10
    \end{bmatrix}
    +
    \begin{bmatrix}
        15 & -44 & -3 \\
        8 & 0 & 9
    \end{bmatrix}
    =
    \begin{bmatrix}
        17 & -38 & 9 \\
        15 & -4 & 19
    \end{bmatrix}
\end{equation*}

\subsection*{Scalar multiplication}

Multiply every number in the matrix by the scale factor.

\begin{equation*}
    2 * 
    \begin{bmatrix}
        4 & 0 \\
        1 & -9
    \end{bmatrix}
    =
    \begin{bmatrix}
        8 & 0 \\
        2 & -18
    \end{bmatrix}
\end{equation*}

\subsection*{Multiplication (dot product)}

To get the number in row i and column j of the resulting matrix, we multiply each number in row i of the first matrix by each number in column j of the second and add the products.

If C = AB, for an $n \times m$ matrix A and an $m \times p$ matrix B, then C is an $n \times p$ matrix. Note that the number of columns in matrix A and the number of rows in matrix B must be the same.

To find A $\cdot$ B:

\lstinputlisting{math/code/matrices1.java}

\subsection*{Exponentiation}

Matrix exponentiation is essentially the same as numerical exponentiation, and can be used with fast-expo techniques. Note that raising a matrix to the zeroth power returns the identity matrix and that the matrix must be square.

\lstinputlisting{math/code/matrices2.java}

\newpage